SiO Molecular Jets around young stars - A numerical perspective

### 2013-08-30 PAPER/ms_final_bv.tex
Added new figs and text

\subsection{Comparison of cooling prescription}
Importance of cooling prescriptions\\
Bulk of emission comes from cooling instabilities.
figure nos. 2 -- Simulated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_90_empvspec.png}
 \caption{Emission map ({\it top left}), PV diagram ({\it top right}
   and spectra ({\it bottom} for the
   2-$>$1 SiO transition for the run with molecular cooling with
   $\eta$ = 3 and fractional abundance of functional form is used
 along with $\delta < $ 1. The jet is assumed to be in the plane of
 sky implying an angle of inclination of $90^{\circ}$. The horizontal
 {\it white dashed} line marks the cuts where the spectra is taken
 while the vertical {\it magenta dashed} line represents the cut for
 the PV digram.} 
\label{empvspec90}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_60_empvspec.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $60^{\circ}$.} 
\label{empvspec60}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_45_empvspec.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $45^{\circ}$.} 
\label{empvspec45}
\end{figure}

@@

Figure~\ref{fig:ltconts} show symmetrical contour maps of integrated 
emission convolved with a 2\arcsec beam coming from different line
transitions for our reference model. The most
striking feature seen in that figure is the progressive shift of
emitting region from the interface to internal knots with increase in
excitation energies of different lines. In particular, the lowest
transition J=1-$>$0 shows most of the emission coming from the
interface between the jet and ambient medium, along with emission from
the dense knots formed at the base of the flow. While, emission coming
due to high energy transitions, J=5-$>$4 and 8-$>$7, 
is more concentrated in the inner jet regions and arise mainly from
the shocks due to internal knots. These high level transitions
also show an evidence of sub-structure close to the primary bow shock.
Such a trend in emission with excitation energies coming from different SiO line transitions
have been observed in many young outflows (for e.g., {\it{L1448}}
bullets \citep{Nisini:2007p13128}, {\it{HH211}}
\citep{Chandler:2001p14376, Nisini:2002p14418,
  Hirano:2006p14411}). 
%

@@

The variation of integrated single dish emission with upper transition
levels J$_{\rm up}$ for different abundance profiles is shown in the
right panel of fig.~\ref{fig:lineratio}. It is clear, that a emission
obtained using a top hat profile is brighter as compared to one
obtained with a gaussian profile, due to its obvious high estimate of
SiO abundance in regions of interest. However, curves obtained from
both these profiles show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. Similar profiles are also predicted and
used to estimate the physical conditions assuming optically
thin emission from LVG modeling for L1448 and L1157
\cite{Nisini:2007p13128}. However, this usual approach did not give
accurate results for line ratios SiO(8-7)/(5-4) and SiO(5-4)/(2-1)
in case of HH212. Values $\approx$ 1 for both ratios were
only achieved in optically thick LTE regime
\citep{Cabrit:2007p13804, Lee:2008p13697}. 
Similar values are also obtained for molecular outflows from massive
young stellar object IRAS 17233–3606\cite{Leurini:2013p13165}.
These line ratios from the present
non-LTE radiative transfer model are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. In summary, the
shape, ratio and peak intensity of the spectra obtained from our model
fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in gas phase via shocks.

@@

\subsection{Caveats and Steps Ahead}
Better handle on Molecular fraction in jets.\\
Zero order approximation of SiO fractional abundance profile.\\ 
Since SiO in gas phase for 0 years with our abundance. While SiO takes
10$^{4}$ years. History of SiO in gas phase and proper chemistry to get
abundance.

Wider CO and H2O WISH survey component.\\

>>>>

\subsection{Comparison of cooling prescription}
\label{ssec:coolres}
Importance of cooling prescriptions\\
Bulk of emission comes from cooling instabilities.
figure nos. 2 -- Simulated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/refrun_21_90_emspecpv.png}
 \caption{Integrated Emission map ({\it top left}), PV diagram ({\it top right}
   and spectra ({\it bottom} for the
   2-$>$1 SiO transition for the reference run. 
   The jet is assumed to be in the plane of sky implying an angle of
   inclination of $90^{\circ}$. The horizontal
 {\it white dashed} line marks the cuts where the spectra is taken
 while the vertical {\it magenta dashed} line represents the cut for
 the PV digram.} 
\label{empvspec90}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/refrun_21_60_emspecpv.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $60^{\circ}$.} 
\label{empvspec60}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/refrun_21_45_emspecpv.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $45^{\circ}$.} 
\label{empvspec45}
\end{figure}

@@

Figure~\ref{fig:ltconts} show symmetrical contour maps of integrated 
emission convolved with a 2\arcsec beam coming from different line
transitions for our reference model. The most
striking feature seen in that figure is the progressive shift of
emitting region from the interface to internal knots with increase in
excitation energies of different lines. In particular, the lowest
transition J=1-$>$0 shows most of the emission coming from the
interface between the jet and ambient medium, along with emission from
the dense knots formed at the base of the flow. While, emission coming
due to high energy transitions, J=5-$>$4 and 8-$>$7, 
is more concentrated in the inner jet regions and arise mainly from
the shocks due to internal knots. Such a trend in emission with excitation energies coming from different SiO line transitions
have been observed in many young outflows (for e.g., {\it{L1448}}
bullets \citep{Nisini:2007p13128}, {\it{HH211}}
\citep{Chandler:2001p14376, Nisini:2002p14418,
  Hirano:2006p14411}). Further, the evolved post-shock gas near the
primary bow shock also show bright emission for these high J transitions. 
This gas is linked to the Rayleigh-Taylor instability associated with cooling flows (see
section~\ref{ssec:coolres}). The sub-structure seen close to the
primary bow shock with high resolution does give a sense of clumpiness
in the flow backing the suggestions to explain clumpy SiO emission by
\citealt{Chandler:2001p14376}.
%

@@

The variation of integrated single dish emission with upper transition
levels J$_{\rm up}$ for different abundance profiles is shown in the
right panel of fig.~\ref{fig:lineratio}. It is clear, that a emission
obtained using a top hat profile is brighter as compared to one
obtained with a gaussian profile, due to its obvious high estimate of
SiO abundance in regions of interest. However, curves obtained from
both these profiles show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. Similar profiles are also predicted and
used to estimate the physical conditions assuming optically
thin emission from LVG modeling for L1448 and L1157
\cite{Nisini:2007p13128}. However, this usual approach did not give
accurate results for line ratios SiO(8-7)/(5-4) and SiO(5-4)/(2-1)
in case of HH212. Values $\approx$ 1 for both ratios were
only achieved in optically thick LTE regime
\citep{Cabrit:2007p13804, Lee:2008p13697}. 
Similar values are also obtained for molecular outflows from massive
young stellar object IRAS 17233-3606 \cite{Leurini:2013p13165}.
These line ratios from the present
non-LTE radiative transfer model are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. Further, we have estimated the optical
depth in regions emitting SiO and find them to be optically thick with
an optical depth $\tau \sim 10$. Thus, assuming optically thin SiO emission
from molecular outflows especially from shocked dense knots would
only provide upper limits. In summary, the
shape, ratio and peak intensity of the spectra obtained from our model
fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in gas phase via shocks.

@@

\subsection{Slow Component and Grain Chemistry}
%
We have presented a formidable model to explain the EHV component seen
in SiO emission from several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in presence of
appropriate cooling. The steady state temperature, density and velocities obtained from dynamical
simulation is post-processed to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in gas phase from grain-grain and grain-mantle
collisions exsist, their focus is mainly on weaker shocks
\cite{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800}. 
The pulsating jet propogation model presented here have shock speeds
reaching upto 100 km\,s$^{-1}$. For simplicity we have used two basic
prescription of SiO abundance profile as a function of shock
velocity. Though the shape of the profile is uncertain the upper and
lower limits used are backed by observational evidences. Ideally for a consistent
depedence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. Further by evolving the dynamical simulations with complex
grain chemistry for more than 10$^{4}$ years (time to destroy SiO in
gas phase \cite{Codella:1999p12584}),  one can trace the history of 
SiO in molecular outflows. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper.  
%

The model presented in this paper target the very early phase of molecular
outflows i.e., dynamical time scale of 1000 years.
As discussed in sections above, the unqiue combination of pulsating
jet with chemistry and cooling followed by non-LTE radiative transfer
calculations has shown success in fiting the line shapes, integrated intensity and line ratios of SiO
transitions arising from EHV gas. However, the synthetic spectra show
no signatures of slow SiO component that accompanies EHV emission seen
in many outflows. The physical mechanism for the origin of slow SiO component is still a matter
of debate. One of the suggestion is it arises due to slowing down of
shocked gas as they age. The time scale estimated for the shocked gas
to slow down is of the same order of magnitude as the SiO destruction
time scale i.e., 10$^{4}$ years \cite{Codella:1999p12584}. In this
case, we will not be able to see the effect as the dynamical runs do
not evolve beyond 10$^{3}$ years. However, we do see initial 
hints of slowing down of more evolved gas in terms of the velocity
shift of about 5-7 km\,s$^{-1}$ in peak emission between that
coming from the freshly formed internal knots close to the base of the
flow as compared to emission coming from more evolved gas near the
primary bow shock. To ascertain this fact in more details, one would
need to track the primary bow shock for 10$^{4}$ years
using a larger simulation box and possibly with adaptive gridding 
which will be considered for future simulations. Such a long term
evolution will also help to provide a numerical model for the formation of HH objects
which are belived to be slowed down and more evolved form of these
young molecular bullets as suggested by \cite{Norman:1979p14858,
  Hartigan:1987p11178}.
%

Alternative to this evolutinary
